---
title: "BMC-Paper"
author: "Daksh Saksena"
date: "February 7th 2023"
---

This is an R markdown notebook replicating the process from a paper on biomarkers in atrial fibrillation from Li-Da Wu et. al. 
Citation: Wu, LD., Li, F., Chen, JY. et al. Analysis of potential genetic biomarkers using machine learning methods and immune infiltration regulatory mechanisms underlying atrial fibrillation.
BMC Med Genomics 15, 64 (2022). https://doi.org/10.1186/s12920-022-01212-0

Before starting, this is my first R markdown file on my local system, so I want to download some packages.

```{r}
setwd("/Users/dakshsaksena/Desktop/Practice-Coder/Biomarker")


```

First, we need to download two datasets (GSE41177 and GSE79768) related to Atrial fibrillation (AF) from GEO, and this can be done through GEOquery. 

```{r}
if (require("BiocManager", quietly = TRUE))
    install.packages("BiocManager")

BiocManager::install("GEOquery")
```

Now, to download the datasets 

```{r}
library(GEOquery)
gds <- getGEO("GPL570")
my_id <- "GSE41177"
gse <- getGEO(my_id)
my_id2 <- "GSE79768"

gse2 <- getGEO(my_id2)

my_id3 <- "GSE14975"
gse3 <- getGEO(my_id3)

```

Lets look at the data first, I found a super helpful video on this called https://www.youtube.com/watch?v=JQ24T9fpXvg
```{r}
## check how many platforms used
length(gse)
gse <-gse[[1]]
gse
pData(gse)$data_processing[1]

## have a look on the expression value
summary(exprs(gse))
```
```{r}
gse2 <- getGEO(my_id2)
length(gse2)
gse2 <-gse2[[1]]
gse2
summary(exprs(gse2))

```
# length(gse3)
# gse3 <-gse3[[1]]
# gse3
# pData(gse3)$data_processing[1]

## have a look on the expression value
#summary(exprs(gse2))

```
The values look well within 0 to 16, but lets still perform log2 transformation since that is what the authors did. 
Now, to perform log2 expression, we do this, then look at the expression value again, and then plot it
```{r}
## have a look on the expression value
summary(exprs(gse))
# summary(exprs(gse2))
# summary(exprs(gse3))
exprs(gse) <- log2(exprs(gse))
summary(exprs(gse))

boxplot(exprs(gse),outline=F)
```

Similarly, for gse2

```{r}
exprs(gse2) <- log2(exprs(gse2))
summary(exprs(gse2))

boxplot(exprs(gse2),outline=F)
```


```{r}
## have a look on the expression value
summary(exprs(gse2))
# summary(exprs(gse2))
# summary(exprs(gse3))
exprs(gse2) <- log2(exprs(gse2))
summary(exprs(gse2))

boxplot(exprs(gse2),outline=F)
```

Both plots look similar, and according to me the log2 expression did not change much, however, it is still imperative.

The next step of the paper is performing data processing and DEGs screening, which is done with the help of an R/bioconductor package called limma. 
After downnlading and calling limma, Wu et. al proceeded to perform Log2-transformation and background correction to the datasets. 
```{r}
BiocManager::install("limama")
library ("limma")
```
Method of background correcting for the purpose of assessing differential expression. Method is taken from Limma documentation
```{r}
gse <- backgroundCorrect(gse, method="normexp", offset=50)
gse2 = backgroundCorrect(gse2, method="normexp", offset=50)
```
Now to merge them

```{r}
comb_gse = combine (gse,gse2)
```

And, now, we have to visualise then
```{r}
install.packages("pheatmap")
install.packages("ggplot2")
 
# load package
library(pheatmap)
library(ggplot2)


#Visualise heatmap and volcano plot
pheatmap(comb_gse)
```

:::
Before performing batch effect adjustment, I wanted to learn what it is exactly. The authors used SVA package to do batch effect adjustment.
From the SVA package, Batch effects are sub-groups of measurements that have qualitatively different behaviour across conditions and are 
unrelated to the biological or scientific variables in a study. 


# Do batch Effect adjustment, as I did not understand how to use SVA package to do 


